Automated Program Verification and Synthesis of Embedded \& Cyber-physical Systems: Advances and Challenges

### 2017-08-03 paper.tex
minor corrections
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

\textbf{(RP4)} State-of-the-art SMT-based context-BMC approaches were extended to verify overflow, limit cycle, time constraints, stability, and minimum phase, in digital systems. Indeed, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter} were tackled, in order to specify system-level properties of those systems, using linear-time temporal logic~\cite{JMorse15}. In particular, a specific UAV application was tackled, with the goal of verifying its attitude controllers~\cite{Bessa16,Chaves17}. More recently, other implementation aspects were added to the existing verification framework, including magnitude, phase, poles, and zeros, which provide deeper analysis regarding digital systems, by tackling frequency-domain parameters, permissible deviation, and associated natural response \cite{sbrtdaniel}. In general, however, there is still a lack of studies to verify system-level properties related to embedded systems; emphasis should be given to micro-grids~\cite{xu15} and cyber-physical systems~\cite{leeCPS2}, which require high-dependability requirements for computation, control, and communication. Additionally, the application of automated fault detection, localization, and correction techniques to digital systems represents an important research direction to make BMC tools useful for embedded systems engineers~\cite{Alves15}.

@@

\textbf{(RP7)} Embedded systems are typically implemented in low- and medium-level programming languages ({\it e.g.}, ANSI-C/C++/Java). Given the availability of different programming languages for writing code for embedded systems, the provision of a unified framework able to tackle low- and high-level properties, as well as system behavior, is needed to achieve code coverage in real applications. The clang compiler~\cite{Lopes:2014:GSL:2692607} seems to offer a simple and strong approach to lead software verification to another degree of completeness so that every implementation aspect of an embedded system can be addressed and evaluated. In particular, by using the clang compiler as front-end for a software verification tool, we avoid the need for maintaining a proprietary front-end (a real challenge nowadays, given that the C and C++ standards are rapidly evolving), and then focus on the main tool objective, the formal program verification. Here, we provide only operation models to tackle low- and high-level system properties via an abstract representation of the associated libraries, which conservatively approximates their semantics. The clang compiler offers an industrial quality analyzer and an application programming interface to access and traverse its internal AST, which can then be used by software verifiers to generate their IR or alternatively analyze the LLVM bitcode produced by clang.

Lastly, yet importantly, BMC tools like CBMC~\cite{Clarke04}, ESBMC~\cite{MorseCNF13,MorseRCN014}, SMACK~\cite{Haran15}, and LLBMC~\cite{MerzFS12} represent the most prominent approaches for verifying C programs, as observed in the Intl. Competition on Software Verification~\cite{Beyer14,BeyerSVCOMP15,Beyer16}, where verifiable (correctness and violation) witnesses are of extreme importance for evaluating software verifiers~\cite{BeyerW15,RochaIFM12}. The increasing number of verification tasks being added every year to this competition allows developers to further improve their verifiers by implementing new types of abstraction and simplification techniques. As a consequence, the continuous development and improvement of verifiers can leverage more efficient synthesizers and further increase the reliability of verified (embedded) systems.

>>>>

\textbf{(RP4)} State-of-the-art SMT-based context-BMC approaches were extended to verify overflow, limit cycle, time constraints, stability, and minimum phase, in digital systems. Indeed, digital filters and controllers~\cite{dsv_spin2015,esbmc_controller,esbmc_filter} were tackled, in order to specify system-level properties of those systems, using linear-time temporal logic~\cite{JMorse15}. In particular, a specific UAV application was tackled, with the goal of verifying its attitude controllers~\cite{Bessa16,Chaves17}. More recently, other implementation aspects were added to the existing verification framework, including magnitude, phase, poles, and zeros, which provide deeper analysis regarding digital systems, by tackling frequency-domain parameters, permissible deviation, and associated natural response \cite{sbrtdaniel}. In general, however, there is still a lack of studies to verify system-level properties related to ECPS; emphasis should be given to micro-grids~\cite{xu15}, which require high-dependability requirements for computation, control, and communication. Additionally, the application of automated fault detection, localization, and correction techniques to digital systems represents an important research direction to make BMC tools useful for ECPS engineers~\cite{Alves15}.

@@

\textbf{(RP7)} Embedded systems are typically implemented in low- and medium-level programming languages ({\it e.g.}, ANSI-C/C++/Java). Given the availability of different programming languages for writing code for ECPS, the provision of a unified framework able to tackle low- and high-level properties, as well as system behavior, is needed to achieve code coverage in real applications. The clang compiler~\cite{Lopes:2014:GSL:2692607} seems to offer a simple and strong approach to lead software verification to another degree of completeness so that every implementation aspect of an embedded system can be addressed and evaluated. In particular, by using the clang compiler as front-end for a software verification tool, we avoid the need for maintaining a proprietary front-end (a real challenge nowadays, given that the C and C++ standards are rapidly evolving), and then focus on the main tool objective, the formal program verification. Here, we provide only operation models to tackle low- and high-level system properties via an abstract representation of the associated libraries, which conservatively approximates their semantics. The clang compiler offers an industrial quality analyzer and an application programming interface to access and traverse its internal AST, which can then be used by software verifiers to generate their IR or alternatively analyze the LLVM bitcode produced by clang.

Lastly, yet importantly, BMC tools like CBMC~\cite{Clarke04}, ESBMC~\cite{MorseCNF13,MorseRCN014}, SMACK~\cite{Haran15}, and LLBMC~\cite{MerzFS12} represent the most prominent approaches for verifying C programs, as observed in the Intl. Competition on Software Verification~\cite{Beyer14,BeyerSVCOMP15,Beyer16}, where verifiable (correctness and violation) witnesses are of extreme importance for evaluating software verifiers~\cite{BeyerW15,RochaIFM12}. The increasing number of verification tasks being added every year to this competition allows developers to further improve their verifiers by implementing new types of abstraction and simplification techniques. As a consequence, the continuous development and improvement of verifiers can leverage more efficient synthesizers and further increase the reliability of verified ECPS.
